Genome-wide changes in expression during inflorescence architecture development

### 2015-09-29 ms.Rmd
Added code to exclude unexpressed ALOG genes but this is disabled for now.

The primary sequences of *Arabidopsis thaliana*, *Solanum lycopersicum* and *Oryza sativa* MADS proteins were downloaded by querying Phytozome 10.3 [@Goodstein:2012ho] for the PANTHER family PTHR11945 ('MADS BOX PROTEIN') via the `biomaRt` interface [@Thomas:2003db; @Durinck:2009ki]. An initial alignment of all proteins was performed with `Clustal Omega` using a full distance matrix (**ref clustalo**). To clean the alignment, poorly conserved regions where greater than `r minpcnongap` percent of sequences had gaps or less than `r minpcident` percent of pairwise comparisons between sequences had identical residues were removed. To recover MADS genes related to the Arabidopsis MIKC clade, a UGPMA tree was estimated from the cleaned alignment and the tree was cut at the lowest height that returned a clade including all type II Arabidopsis MADS proteins listed in the Arabidopsis transcription factor database on the Arabidopsis Gene Regulatory Information Server (**ref for AGRIS**), with as few type I proteins as possible. Protein sequences in this clade that were longer than `r minProtLength` amino acids were realigned using `Clustal Omega`, and the alignment was cleaned as described above and used to estimate a neighbour-joining tree with the `bionj` algorithm implemented in the `ape` package (**ref:ape***) rooted to a type I MADS sequence (`r og`). L~2~FC change expression values for tomato and Arabidopsis genes were estimated from samples that have previously been described [@Mantegazza:2014di; @Park:2012dg] using the same method described above for the rice samples. The tree was plotted using the `ggtree` extension for `ggplot2` (**ref ggtree**).

@@

```{r f_alogFamily, fig.width = 6.614}
print(f_alogFamily)
```

**`r I(pasteLabel("Figure ", figCount, "f_mfuzzClusters"))`**. Expression of the rice G1 (ALOG) family in inflorescence meristems. Points for are plotted in red if the gene was detected in that sample or in blue otherwise. Three G1 genes, *G1L1*, *G1L2* and *G1L5* (*TAWAWA*) are expressed, and they share a similar pattern of expression.

@@

**`r I(pasteLabel("Figure ", figCount, "f_madsTree"))`**. Neighbour-joining similarity tree of putatative MIKC proteins in *A. thaliana*, *Solanum lycopersicum* and *Oryza sativa* ssp. *japonica*. Named proteins starting with AT_ or OS_ are from Arabidopsis or rice respectively, with synonyms in parentheses, and other proteins are labelled with their unique identifiers. The colour of the label represents the Log~2~-fold change (L~2~FC) in expression of the gene encoding that protein between undifferentiated and differentiated meristems. A white background indicates that no results were returned by `DESeq2` for the gene in question, *e.g.* because of low read counts or an extreme count outlier. For rice, L~2~FC was estimated from the RM and SM samples reported here. For *A. thaliana*, L~2~FC was estimated using the inflorescence meristem and floret meristem samples reported by @Mantegazza:2014di. The sympodial inflorescence meristem and floret meristem samples generated by @Park:2012dg were used to estimate L~2~FC values for tomato genes. Clades are labelled following the nomenclature presented by (**ref: Arora paper**).

@@

```{r sessionInfo, results = "asis"}
pander::pander(sessionInfo())
git <- c(paste("**git branch:**",system("git rev-parse --abbrev-ref HEAD", intern = TRUE)),
          paste("**git hash:**", system("git rev-parse HEAD", intern = TRUE)))
cat("\n")
for (x in git){cat(x, "\n")}
```

>>>>

The primary sequences of *Arabidopsis thaliana*, *Solanum lycopersicum* and *Oryza sativa* MADS proteins were downloaded by querying Phytozome 10.3 [@Goodstein:2012ho] for the PANTHER family PTHR11945 ('MADS BOX PROTEIN') via the `biomaRt` interface [@Thomas:2003db; @Durinck:2009ki]. An initial alignment of all proteins was performed with `Clustal Omega` using a full distance matrix (**ref clustalo**). To clean the alignment, poorly conserved regions where greater than `r minpcnongap` percent of sequences had gaps or less than `r minpcident` percent of pairwise comparisons between sequences had identical residues were removed. To recover MADS genes related to the Arabidopsis MIKC clade, a UGPMA tree was estimated from the cleaned alignment and the tree was cut at the lowest height that returned a clade including all type II Arabidopsis MADS proteins listed in the Arabidopsis transcription factor database on the Arabidopsis Gene Regulatory Information Server (**ref for AGRIS**), with as few type I proteins as possible. Protein sequences in this clade that were longer than `r minProtLength` amino acids were realigned using `Clustal Omega`, and the alignment was cleaned as described above and used to estimate a neighbour-joining tree with the `bionj` algorithm implemented in the `ape` package (**ref:ape***) rooted to type I MADS sequences (`r og`). L~2~FC change expression values for tomato and Arabidopsis genes were estimated from samples that have previously been described [@Mantegazza:2014di; @Park:2012dg] using the same method described above for the rice samples. The tree was plotted using the `ggtree` extension for `ggplot2` (**ref ggtree**).

@@

```{r f_alogFamily, fig.width = 3.150, fig.height = 3.150 * 2}
print(f_alogFamily)
```

**`r I(pasteLabel("Figure ", figCount, "f_alogFamily"))`**. Expression of rice G1/ALOG family genes in inflorescence meristems. Genes were retrieved by querying Phytozome 10.3 for the Pfam domain PF04852 [@Goodstein:2012ho] (**ref: Pfam**). Red or blue points indicate whether the gene was detected or not detected respectively for each library. Only three G1 genes (*G1L1*, *G1L2* and *G1L5*/*TAWAWA*) were detected, but they share a similar pattern of expression.

@@

**`r I(pasteLabel("Figure ", figCount, "f_madsTree"))`**. Neighbour-joining similarity tree of putatative MIKC proteins in *A. thaliana*, *Solanum lycopersicum* and *Oryza sativa* ssp. *japonica*. Named proteins starting with AT_ or OS_ are from Arabidopsis or rice respectively, with synonyms in parentheses, and other proteins are labelled with their unique identifiers. The colour of the label represents the Log~2~-fold change (L~2~FC) in expression of the gene encoding that protein between undifferentiated and differentiated meristems. Where there is no coloured label, the gene for the protein in question was removed from the differential expression analysis by `DESeq2`, which may occur because of low read counts, an extreme count outlier or the gene failing independent filtering [@Love:2014do]. For rice, L~2~FC was estimated from the RM and SM samples reported here. For *A. thaliana*, L~2~FC was estimated using the inflorescence meristem and floret meristem samples reported by @Mantegazza:2014di. The sympodial inflorescence meristem and floret meristem samples generated by @Park:2012dg were used to estimate L~2~FC values for tomato genes. Clades are labelled following the nomenclature presented by (**ref: Arora paper**).

@@

```{r sessionInfo, results = "asis"}
cat(paste0("**branch:** _", system("git rev-parse --abbrev-ref HEAD", intern = TRUE), "_\n\n"))
cat(paste0("**hash:** _", system("git rev-parse HEAD", intern = TRUE), "_\n\n"))
pander::pander(sessionInfo())
```
